---
title: "Visual Assignment"
description: |
  A short description of the post.
author:
  - name: Liu Manye
    url: {}
date: 07-13-2021
output:
  distill::distill_article:
    self_contained: false
---

# Mini Challenge 2 Background
In the roughly twenty years that Tethys-based GAStech has been operating a natural gas production site in the island country of Kronos, it has produced remarkable profits and developed strong relationships with the government of Kronos. However, GAStech has not been as successful in demonstrating environmental stewardship.

In January, 2014, the leaders of GAStech are celebrating their new-found fortune as a result of the initial public offering of their very successful company. In the midst of this celebration, several employees of GAStech go missing. An organization known as the Protectors of Kronos (POK) is suspected in the disappearance, but things may not be what they seem.

As an expert in visual analytics, you are called in to help law enforcement from Kronos and Tethys.

In Mini-Challenge 2, you are asked to analyze movement and tracking data. GAStech provides many of their employees with company cars for their personal and professional use, but unbeknownst to the employees, the cars are equipped with GPS tracking devices. You are given tracking data for the two weeks leading up to the disappearance, as well as credit card transactions and loyalty card usage data. From this data, I will try to solve the given questions, identify anomalies and suspicious behaviors, and identify which people use which credit and loyalty cards. 

# Question 1
Using just the credit and loyalty card data, identify the most popular locations, and when they are popular. What anomalies do you see? What corrections would you recommend to correct these anomalies? Please limit your answer to 8 images and 300 words.


## R setup
R is the only tools will be used in this assignment.

Run the code below to complete R option setup:
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina=3,
                      echo = TRUE,
                      eva1 = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

Install the needed package for this assignment. This code chunk checks if required packages are installed. If they are not installed, the next line of code will install them. The following line is then use to import the library into the current working environment.
```{r}
packages = c('igraph', 'tidygraph', 'ggraph', 'visNetwork', 'lubridate', 'clock', 'tidyverse', 'dplyr','raster', 'sf', 'tmap', 'gifski', 'mapview')

for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p, character.only = T)
}
```

## Import Dataset

**cc_data.csv** and **loyalty_data.csv** are the two datasets required to answer Question 1. I used read_csv to import them in the code chunk below:
```{r}
cc <- read_csv("MC2/cc_data.csv")
loyalty <- read_csv("MC2/loyalty_data.csv")
glimpse(cc)
glimpse(loyalty)
```

## Data Preparation
According to the glimpse of the cc data and loyalty data, we can see that timestamps in both file are in a character formate. 
Run the following code to change datatype of timestamps into a correct datetime format:
```{r}
cc$timestamp <- date_time_parse(cc$timestamp,
                zone = "",
                format = "%m/%d/%Y %H:%M")
glimpse(cc)

```

There are only date information in the loyalty dataset, hence its format setting is "**%m/%d/%Y**" instead of **%m/%d/%Y %H:%M**
```{r}
loyalty$timestamp <- date_time_parse(loyalty$timestamp,
                zone = "",
                format = "%m/%d/%Y")
glimpse(loyalty)

```

Both table contains date information of GAStech staff's spending behaviour. In order to combine these two seperated table into one, we need to seperate date information in the cc data and create a link between them. Code chunk below will help us to do that.
```{r}
cc$Date <- format(cc$timestamp, format="%Y-%m-%d")
cc$Date <- date_time_parse(cc$Date,
                           zone = "",
                           format = "%Y-%m-%d")
glimpse(cc)

```

After creating the sharing column both table, a full_join verb is used to join cc data and loyalty data together. In the new table, both cases in cc and loyalty will be kept no matter whether an exact match existed or not.
```{r}
card_joined <- cc %>%
  full_join(loyalty, by = c("Date" = "timestamp", "location", "price"))

```
In card_joined, we can get a clearer image of each credit card holders' spending patterns. Including the timestamp, location, price spent, credit card number, and loyal card number.
In this way, we can see all the existing credit_card and loyalty_card combinations and explore if there is any cross usage of staff's loyalty card in the past half month.


filter popular locations for credit card transactions: 
Code chunk below help use count the number of occurrences of each location and sort them in a descending order. 
```{r}
popularLoc_cc <- cc %>%
  group_by(location) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

```

Similar as the previous code chunk, this code chunk will help use to find the popular locations based staffs' loyalty card usage frequency:
```{r}
popularLoc_loyalty <- loyalty %>%
  group_by(location) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

```

PopularLc_cc file contains one more observations than popularLoc_loyalty, which is Daily Dealz. Daily Dealz only appear once, hence it would not impact on our exploration of Q1 answers.
```{r}
popularLoc_cc %>%
  anti_join(popularLoc_loyalty, by = c("location"))

```


## Charting and Analysis

### Popular Locations

To illustrate the popular locations that GasTech staffs shop frequently, two bar plots will be drawn below.

Running the code chunk below will create a table that contains the top 6 popular location based on credit card usage information so that we can use it to draw a graph:
```{r}
popular_top6_credit <- popularLoc_cc %>%
 
 gather(location, count) %>%
 arrange(desc(count)) %>%
 top_n(6)

popular_top6_credit

```
Run the code chunk below to draw the bar chart:
```{r}
top6Loc_credit<-ggplot(data=popular_top6_credit, aes(x=location, y=count)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=count), position=position_dodge(width=0.9), vjust=-0.25) +
  theme_minimal()

print(top6Loc_credit + ggtitle("Top 6 Popular Locations Based on Credit Card U`sage"))

```

Running the code chunk below to create a table that contains the top 6 popular location based on loyalty card use frequency so that we can use it to draw a graph:
```{r}
popular_top6_loyalty <- popularLoc_loyalty %>%
 
 gather(location, count) %>%
 arrange(desc(count)) %>%
 top_n(6)

popular_top6_loyalty

```

Run the code chunk below to draw the bar chart for popular_top6_loyalty table:
```{r}
top6Loc_loyalty<-ggplot(data=popular_top6_loyalty, aes(x=location, y=count)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=count), position=position_dodge(width=0.9), vjust=-0.25) +
  theme_minimal()

print(top6Loc_loyalty + ggtitle("Top 6 Popular Locations Based on Loyalty Card Usage"))

```


Two popularLoc data shares same top six popular locations, they are: "Katerina's Cafe", "Hippokampos", "Guy's Gyros", "Brew've Been Served", "Ouzeri Elian", "Hallowed Grounds". There is a small different in the ranking of these location in two tables but it is insignificant for our exploration.

### Popular Time Periods

Create a table contains popular location info only:
```{r}
popular_locations <- card_joined %>%
  filter(location %in% c("Katerina's Cafe", "Hippokampos", "Guy's Gyros", "Brew've Been Served", "Ouzeri Elian", "Hallowed Grounds")) %>%
  drop_na(timestamp) %>%
  dplyr::select(-Date)

```
According to this table, we can see that Katrina's Cafe, Hippokampos, Guy's Gyros, and Ouzeri Elian are places that sells food, hence these location are popular during the lunch break (12-14 pm.) and dinner time (19-21 pm.).
On the other hand, Brew've Been Served and Hallowed Grounds mainly sell coffee. They are popular between 7:30 - 8:30 am. only since GasTech staffs only visit there to grab a coffee before working.

### Anomalies

To promote local businesses, Kronos based companies provide a Kronos Kares benefit card to GASTech employees giving them discounts and rewards in exchange for collecting information about their credit card purchases and preferences as recorded on loyalty cards. 
Since loyalty card is discount card that offered by GasTech is meant to promote business and to collect staffs' spending information, it is abnormal to see cross usage of loyalty card between GasTech staffs. Anomolies like this implies that there is a possible hidden relationship between the owners of loyalty cards that were cross used.

Run the code chunk below to find out credit cards that involves in the anomalies discussed above:
```{r}
abnormal_cc <- popular_locations %>%
  drop_na(loyaltynum) %>%
  group_by(last4ccnum) %>%
  summarize(loy_n = n_distinct(loyaltynum)) %>%
  filter(loy_n > 1)

abnormal_cc
```

# Question 2

Add the vehicle data to your analysis of the credit and loyalty card data. How does your assessment of the anomalies in question 1 change based on this new data? What discrepancies between vehicle, credit, and loyalty card data do you find? Please limit your answer to 8 images and 500 words.

## Developing path track map 

Importing raster file MC2-tourist.tif into R by using raster() or Raster package.
```{r}
bgmap <- raster("MC2/MC2-tourist.tif")
bgmap

```

In the code chunk below, we will use tm_raster() to plot a raster layer by using tmap package.
```{r}
tmap_mode("plot")
tm_shape(bgmap) +
  tm_raster(bgmap,
            legend.show = FALSE)

```

```{r}
tm_shape(bgmap) +
tm_rgb(bgmap, r = 1,g = 2,b = 3,
       alpha = NA,
       saturation = 1,
       interpolate = TRUE,
       max.value = 255)

```

Importing Vector GIS Data File:
Abila GIS data layer is in ESRI shapefile format. Code-chunk below will use st_read() of the sf package to import Abila shapefile into R.
```{r}
Abila_st <- st_read(dsn = "MC2/Geospatial",
                    layer = "Abila")
                    
```

Using read_csv() of readr package to import gps2.csv into R.
```{r}
gps2 <- read_csv("MC2/gps2.csv")
glimpse(gps2)

```

New factor **Time Difference** was previously added into the gps file by using the Timestampe of the next gps record of a particular car id minus the previous Timestampe. After calculating the time differences between each gps records of a car id, present it in a second unit to get **Seconds**.
In this way, we can get the time interval between a person's movement and using it to predict its spending behaviour in the future exploration.

change datatypes
```{r}
gps2$Timestamp <- date_time_parse(gps2$Timestamp,
                zone = "",
                format = "%m/%d/%Y %H:%M")
gps2$id <- as_factor(gps2$id)

glimpse(gps2)

```

```{r}
gps_sf <- st_as_sf(gps2, 
                   coords = c("long", "lat"),
                       crs= 4326)
```

```{r}
gps_sf$day <- format(gps_sf$Timestamp, format="%d")
gps_sf$hour <- format(gps_sf$Timestamp, format="%H")
gps_sf$minute <- format(gps_sf$Timestamp, format="%M")
```

#find out time interval > 5mins
```{r}
more_than_3mins <- gps_sf %>%
  filter(Seconds > 180)
```


Creating Movement Path from GPS Points
```{r}
gps_path <- gps_sf %>%
  group_by(id, day, hour) %>%
  summarize(m = mean(Timestamp), 
            do_union=FALSE) %>%
  st_cast("LINESTRING")
```


```{r}
p = npts(gps_path, by_feature = TRUE)
gps_path2 <- cbind(gps_path, p) %>%
  filter(p>1)

```

# Drawing line graphs
```{r}
gps_path_selected <- gps_path2 %>%
  filter(id==29, hour == "20")
tmap_mode("view")
tm_shape(bgmap) +
  tm_rgb(bgmap, r = 1,g = 2,b = 3,
       alpha = NA,
       saturation = 1,
       interpolate = TRUE,
       max.value = 255) +
  tm_shape(gps_path_selected) +
  tm_lines()
```


```{r}
gps_dot <- more_than_3mins %>%
  group_by(id, hour, day, minute) %>%
  summarize(geo_n = n_distinct(geometry)) %>%
  st_cast("POINT")
```

# Drawing dot graph
```{r echo=FALSE, eval=TRUE, fig.height=6}
gps_dots_selected <- gps_dot %>%
  filter(day=='06',hour=='13')
tmap_mode("view")
tm_shape(bgmap) +
  tm_rgb(bgmap, r = 1,g = 2,b = 3,
       alpha = NA,
       saturation = 1,
       interpolate = TRUE,
       max.value = 255) +
  tm_shape(gps_dots_selected) +
  tm_dots()
```


read data

```{r}
car <- read_csv("MC2/car-assignments.csv")
glimpse(car)

```

Change datatype
```{r}
car <- car %>%
  drop_na(CarID)

car$CarID <- as_factor(car$CarID)

glimpse(car)
glimpse(gps2)

```

join cc data and loyalty data together
```{r}
car_gps <- car %>%
  full_join(gps2, by = c("CarID" = "id"))

```

edit columns
```{r}
car_gps <- st_as_sf(car_gps, 
                   coords = c("long", "lat"),
                       crs= 4326)

car_gps <- car_gps %>%
  unite("Name", FirstName, LastName, sep = " ")
```














